### 2014-09-02 constraints.tex
changed structure

\section{Constraint Derivation and Resolution}

We now describe in detail the constraint encoding and solving steps.

\subsection{Same Trace}

We begin with the basic setting, which enforces conservative reordering of the events along the input trace without diverging into new dependence structures or execution paths. We relax this requirement later, such that different data-flow paths and branches can be followed as long as feasibility is retained.

\paragraph{Intra-thread Order}

The first set of constraints reflects control flow within the individual threads, which must remain unchanged under the reordering transformation. Given input trace $t$, this is expressed as the following formula:
$$
\begin{array}{rl}
	\forall \tau,\tau' \in t. & {\sf thread}\ \tau \equiv {\sf thread}\ \tau'. \\
										& {\sf index}\ \tau < {\sf index}\ \tau' \Rightarrow O_{\tau} < O_{\tau'}
\end{array}
$$ 
The logical variables $O_x$ express ordering constraints. The requirement, as stated above, is that these variables
reflect the same order as the projection of ${\sf index}$ onto individual threads.

\paragraph{Race Condition}

Given pair $\tau$ and $\tau'$ of events that both access a common memory location $\ell$, we demand that
$$
\begin{array}{rl}
					& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
\bigwedge 	& (\ell \in {\sf writeset}\ \tau \vee \ell \in {\sf writeset}\ \tau') \\
\bigwedge   & O_{\tau} = O_{\tau'}
\end{array}
$$
That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$ (as judged by the ${\sf writeset}$ membership check), and (iii) the events occur simultanesouly. This is a direct logical encoding of the definition of a race condition.

\paragraph{Path Constraints}

The requirement with respect to branching is that the prefix of the original trace $t$ up to the pair $\tau$ and $\tau'$ of candidate racing events remains identical in the predicted trace $t'$. More accurately, under the assumption that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
we require that 
$$
	\bigwedge_{\tau'' \in t \cap {\bf bexp}.\
		\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
$$ 
where ${\sf stmt}$ is a helper function that obtains the code statement incident in a given transition. That is, all branching transitions up to $\tau'$ (which occurs after $\tau$) preserve their boolean interpretation under $t'$. This ensures that there are no divergences from the path containing the racing events, though beyond that path any feasible continuation is permitted. Importantly, contrary to \cite{JEFF-PLDI14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions is invariant under the input and predicted traces.

\paragraph{Variable Definitions}

A final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before. That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$. This is enforced as the formula
$$
	\bigwedge_{\tau'' \in t \cap {\bf asgn}.\
		\tau'' \in {\sf pre}\ \tau'} {\sf stmt}\ \tau'' \in t' 
$$
where we again assume that ${\sf index}\ \tau < {\sf index}\ \tau'$. That is, the same statement occurring in $t$ is also present in $t'$ (though the transitions may differ). Since the statements of $t'$ are a permutation of the statements of $t$, we are assured that use/def flow is constrained appropriately.

\subsection{Relaxations}

We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the data flow exhibited in the original trace. More precisely, \tool\ is able to relax flow dependencies in the original trace, whereby a thread reads a shared memory location written by another thread, while enforcing feasibility. This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same write access as in the original trace.

\paragraph{Relaxation of Flow Dependencies}

To ensure feasibility under relaxation of flow dependencies, we need to secure the link between the execution schedule and the write/read flow. As an illustration from the example in Figure \ref{fig:running}, $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule where thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}, the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}: $R_{\tt y}^4={\tt y}^1$.

The full and general constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
$$
\begin{array}{rll}
\bigvee_{e_w \in {\cal W}} &  & (R_{\ell} = {\ell}^{{\sf index}\ e_w}) \\
&		\bigwedge 	&  O(e_w) < O(e) \\
&		\bigwedge_{e' \in {\cal W} \setminus \{ e_w \}} & (O(e') < O(e_w) \vee O(e) < O(e'))
\end{array}
$$
This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O(e_w) < O(e)$) and (ii) all other write events either occur before ($O(e') < O(e_w)$) it or after the read event
($O(e) < O(e')$).

An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning. As an illustration, we refer to Figure \ref{fig:heapAccess}. While the read at line {\tt 7} appears to match the write at line {\tt 5}, this is conditioned on the read at line {\tt 6} being linked to the assignment at line {\tt 4}. If the predicted run violates this link, then feasibility is no longer guaranteed. In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the write at line {\tt 5} no longer matches the read at line {\tt 7}.

\begin{figure}
	\centering
	\begin{tabular}{ll}
		\hline
		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
		\hline
		{\tt 1: x1 = new();} & \\
		{\tt 2: x2 = new();} & \\
		{\tt 3: y = x1;} & \\
		{\tt 4: y = x2;} & \\
		{\tt 5: x2.f = 5;} & \\	
		& {\tt 6: z = y;} \\
		& {\tt 7: w = z.f;} \\
	\end{tabular}
	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation}
\end{figure}

To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced before the candidate racing events retain their original address in the predicted trace.
This achieves three guarantees: First, matching events in the original trace are guaranteed to also match in the predicted trace. Second, candidate races in the original trace remain viable in the predicted trace. Finally, in a practical setting involving virtual method calls (which goes beyond our core grammar in Table \ref{Ta:syntax}), call-site resolutions are the same across the original and predicted traces. Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.

Formally, given pair $\tau$ and $\tau'$ of candidate racing events such that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
we require that
$$
\bigwedge_{\tau'' \in t \cap {\bf heapr}.\
	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
$$ 
That is, all heap dereferences up to the later of the candidate racing events must retain their original meaning in the predicted trace $t'$. This ensures in particular that the base reference remains identical to the one in the original trace $t$.

@@

\subsection{Proof of Correctness}

>>>>

\section{Exploration of Unexecuted Branches}

We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.

@@

\paragraph{Proof of Correctness}

### 2014-09-02 overview.tex
changed structure

>>>>

\subsection{Preliminaries}

Throughout this paper, we assume a standard operational semantics, which defines (i) a mapping ${\sf thr}$ between
execution threads, each having a unique identifier $i \in \mathbb{N}$, and their respective code, as well as (ii) per-thread stack and shared heap memory.

The code executed by a given thread follows the syntax in Table \ref{Ta:syntax}. The text of a program is a sequence of zero or more method declarations, as given in the definition of symbol {\tt p}. Methods accept zero or more arguments $\overline{{\tt x}}$, have a body ${\tt s}$, and may have a return value (which we leave implicit). For simplicity, we avoid from static typing as well as virtual methods. The body of a method consists of the core grammar for symbol {\tt s}. We avoid from specifying syntax checking rules, as the grammar is fully standard.

For simplicity, we assume that in the starting state each thread points to a parameter-free method. In this way, we can simply assume an empty starting state (i.e., an empty heap), and eliminate complexities such as user-provided inputs and initialization of arguments with default values. These extensions are of course possible, and in practice we handle these cases, but reflecting them in the formalism would result in needless complications.

As is standard, we assume an interleaved semantics of concurrency. A \emph{transition}, or \emph{event}, is of the form
$\sigma \stackrel{i / {\tt s}}{\longrightarrow} \sigma'$, denoting that thread $i$ took an evaluation step
in prestate $\sigma$, wherein atomic
statement {\tt s} was executed, which resulted in poststate $\sigma'$. We refer to a sequence of transitions from
the starting state to either an exceptional state (e.g., due to null dereference) or a state where all the threads have 
reduced their respective code to $\epsilon$ as a \emph{trace}.

\begin{table}
	\begin{center}
		\begin{tabular}{rl}
			\multicolumn{1}{l}{{\tt p} $::=$} & \\
			$\overline{\texttt{m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (mdcl)} \\
			\multicolumn{1}{l}{{\tt s} $::=$} & \\
			{\tt x = y}\ $|$\ {\tt x = $c$} $|$\ & {\bf (asgn)} \\
			{\tt x = new()} $|$\ & {\bf (alloc)} \\
			{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (call)} \\
			{\tt return x} $|$\ & {\bf (return)} \\
			{\tt z = x \{ $+$,$-$,$\times$,$/$ \} y}\ $|$\ & {\bf (aexp)} \\
			{\tt b = x \{ $<$,$\leq$,$\equiv$,$\geq$,$>$ \} y}\ $|$\ & \\
			{\tt b = bx $\wedge$ by}\ $|$\ {\tt b = bx $\vee$ by} $|$\ & {\bf (bexp)} \\
			{\tt y = x.f} $|$\ & {\bf (heapr)} \\ 
			{\tt x.f = y}\ $|$\ {\tt x.f = $c$}\ $|$\ & {\bf (heapw)} \\
			{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (ctrl)} \\
			{\tt lock(x)}\ $|$\ {\tt unlock(x)} & {\bf (sync)}
		\end{tabular}
	\end{center}
	\caption{\label{Ta:syntax}Language syntax}
\end{table}

We make use of the following helper functions:
\begin{itemize}
	\item ${\sf proj}\ t\ i$ projects trace $t$ onto all transitions involving thread $i$.
	\item $t[k]$ obtains the $k$th transition within trace $t$.
	\item ${\sf index}\ t\ \tau$ retrieves the index, or offset, of transition $\tau$ within trace $t$. When simply writing
	${\sf index}\ \tau$ (while omitting the trace parameter) we refer to the index of $\tau$ within the original trace. 
	\item ${\sf pre}\ t\ \tau$ is the prefix of trace $t$ preceding transition $\tau$. For the suffix beyond $\tau$, we 
	use ${\sf post}\ t\ \tau$. Finally, ${\sf bet}\ t\ \tau_1\ \tau_2$ returns the transitions delimited by $\tau_1$ and $\tau_2$.
\end{itemize}

\subsection{Constraint System}

### 2014-09-02 trace.tex
changed structure

\section{Trace Encoding}

@@

In this section, we describe how a concrete execution trace is encoded as a symbolic trace, which enables \tool\ to 
derive reordering constraints.

\subsection{Preliminaries}

Throughout this paper, we assume a standard operational semantics, which defines (i) a mapping ${\sf thr}$ between
execution threads, each having a unique identifier $i \in \mathbb{N}$, and their respective code, as well as (ii) per-thread stack and shared heap memory.

The code executed by a given thread follows the syntax in Table \ref{Ta:syntax}. The text of a program is a sequence of zero or more method declarations, as given in the definition of symbol {\tt p}. Methods accept zero or more arguments $\overline{{\tt x}}$, have a body ${\tt s}$, and may have a return value (which we leave implicit). For simplicity, we avoid from static typing as well as virtual methods. The body of a method consists of the core grammar for symbol {\tt s}. We avoid from specifying syntax checking rules, as the grammar is fully standard.

For simplicity, we assume that in the starting state each thread points to a parameter-free method. In this way, we can simply assume an empty starting state (i.e., an empty heap), and eliminate complexities such as user-provided inputs and initialization of arguments with default values. These extensions are of course possible, and in practice we handle these cases, but reflecting them in the formalism would result in needless complications.

As is standard, we assume an interleaved semantics of concurrency. A \emph{transition}, or \emph{event}, is of the form
$\sigma \stackrel{i / {\tt s}}{\longrightarrow} \sigma'$, denoting that thread $i$ took an evaluation step
in prestate $\sigma$, wherein atomic
statement {\tt s} was executed, which resulted in poststate $\sigma'$. We refer to a sequence of transitions from
the starting state to either an exceptional state (e.g., due to null dereference) or a state where all the threads have 
reduced their respective code to $\epsilon$ as a \emph{trace}.

\begin{table}
	\begin{center}
	\begin{tabular}{rl}
						\multicolumn{1}{l}{{\tt p} $::=$} & \\
					    $\overline{\texttt{m($\overline{{\tt x}}$) \{ s \}}}$ & {\bf (mdcl)} \\
						\multicolumn{1}{l}{{\tt s} $::=$} & \\
						{\tt x = y}\ $|$\ {\tt x = $c$} $|$\ & {\bf (asgn)} \\
						{\tt x = new()} $|$\ & {\bf (alloc)} \\
						{\tt x = f($\overline{{\tt y}}$)} $|$\ & {\bf (call)} \\
						{\tt return x} $|$\ & {\bf (return)} \\
						{\tt z = x \{ $+$,$-$,$\times$,$/$ \} y}\ $|$\ & {\bf (aexp)} \\
						{\tt b = x \{ $<$,$\leq$,$\equiv$,$\geq$,$>$ \} y}\ $|$\ & \\
						{\tt b = bx $\wedge$ by}\ $|$\ {\tt b = bx $\vee$ by} $|$\ & {\bf (bexp)} \\
						{\tt y = x.f} $|$\ {\tt x.f = y}\ $|$\ & {\bf (heap)} \\
						{\tt if (b) \{ s \} $|$\ {\tt while (b) \{  s \}} $|$\ {\tt s ; s }} $|$\ & {\bf (ctrl)} \\
						{\tt lock(x)}\ $|$\ {\tt unlock(x)} & {\bf (sync)}
	\end{tabular}
	\end{center}
	\caption{\label{Ta:syntax}Language syntax}
\end{table}

We make use of the following helper functions:
\begin{itemize}
	\item ${\sf proj}\ t\ i$ projects trace $t$ onto all transitions involving thread $i$.
	\item $t[k]$ obtains the $k$th transition within trace $t$.
	\item ${\sf index}\ t\ \tau$ retrieves the index, or offset, of transition $\tau$ within trace $t$. When simply writing
	 ${\sf index}\ \tau$ (while omitting the trace parameter) we refer to the index of $\tau$ within the original trace. 
	\item ${\sf pre}\ t\ \tau$ is the prefix of trace $t$ preceding transition $\tau$. For the suffix beyond $\tau$, we 
	use ${\sf post}\ t\ \tau$. Finally, ${\sf bet}\ t\ \tau_1\ \tau_2$ returns the transitions delimited by $\tau_1$ and $\tau_2$.
\end{itemize}

\subsection{Basic Encoding: Local Accesses}

We describe how \tool\ encodes a trace into symbolic form in two steps. We first start with ``local'' intraprocedural events --- i.e., 
transitions that manipulate memory accessed by at most a single thread during an invocation-free run --- and then extend our encoding scheme to the entire set of possible events.

\paragraph{SSA Form} The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
a variable is defined exactly once. In this way, def/use chains become explicit, and encoding of trace events as constraints is simplified. As an illustration, trace
\begin{quote}
	{\tt 1: x=1; 2: x<3; 3: x=3;} \\
	{\tt 4: y=1;} \\
	{\tt 5: z=x+y}
\end{quote}
becomes
\begin{quote}
	{\tt 1: x$^1$=1; 2: x$^1$<3; 3: x$^3$=3;} \\
	{\tt 4: y$^4$=1;} \\
	{\tt 5: z$^5$=x$^3$+y$^4$}
\end{quote}
(For readability, we version variables according to the line number of their definition.)

\paragraph{Local Heap Accesses} Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses. While in general determining whether a given heap access is local is an undecidable problem, in the context
of a concrete execution trace this determination is straightforward. We can thus improve upon the baseline SSA form, where def/use chains over heap accesses are ignored. Indeed, to ensure feasibility, we must account for such accesses. Hence, we replace accesses to field {\tt f} of local object $o$ with a fresh local variable {\tt l$_{o{\tt .f}}$}, where $o{\tt .f}$ denotes the memory location itself and not its value. This is done prior to the standard SSA transformation, and under the assumption that in the predicted run, the base object $o$ remains the same.

\subsection{Full Encoding: Method Calls and Shared Accesses}

@@

\paragraph{Method Invocations} To account for method invocations, we induce additional context on variable accesses. Instead of recording only the version of a variable, we also represent as part of the variable's identifier its enclosing method and its invocation counter. Thus, symbol ${\tt m}^2 \colon {\tt v}^7$ is interpreted as the definition of local variable ${\tt v}$ within the second invocation of method ${\tt m}$ that occurs at trace index 7.

@@

\paragraph{Shared Heap Accesses} The final aspect of our encoding process is representation of shared heap accesses. As illustrated in Section \ref{Se:techoverview}, this is done via designated symbols:
\begin{itemize}
	\item $W^{k}_{\tt v}$ denotes write access to variable ${\tt v}$ at trace index $k$ (where ${\tt v}$ is assumed to be a qualified identifier to account for the interprocedural setting).
	\item Analogously, $R^{k}_{\tt v}$ denotes that ${\tt v}$ is read at index $k$.
\end{itemize}

@@

>>>>

\section{Relaxation of Flow Dependencies}

@@

We begin by describing the first relaxation, whereby the dependence structure defined by the original trace is potentially violated while retaining soundness.

\subsection{Trace Encoding}

The first step of the process is to rewrite the concrete execution trace into a symbolic form that enables \tool\ to 
derive reordering constraints. We describe how \tool\ encodes a trace into symbolic form in two steps. We first start with ``local'' intraprocedural events --- i.e., 
transitions that manipulate memory accessed by at most a single thread during an invocation-free run --- and then extend our encoding scheme to the entire set of possible events.

\paragraph{Basic Encoding: Local Accesses}

The fundamental encoding transformation is to induce Static Single Assignment (SSA) form on the raw trace, such that
a variable is defined exactly once. In this way, def/use chains become explicit, and encoding of trace events as constraints is simplified. As an illustration, trace
\begin{quote}
	{\tt 1: x=1; 2: x<3; 3: x=3;} \\
	{\tt 4: y=1;} \\
	{\tt 5: z=x+y}
\end{quote}
becomes
\begin{quote}
	{\tt 1: x$^1$=1; 2: x$^1$<3; 3: x$^3$=3;} \\
	{\tt 4: y$^4$=1;} \\
	{\tt 5: z$^5$=x$^3$+y$^4$}
\end{quote}
(For readability, we version variables according to the line number of their definition.)

Beyond the default SSA rewriting algorithm, we apply a specialized transformation to handle local heap accesses. While in general determining whether a given heap access is local is an undecidable problem, in the context
of a concrete execution trace this determination is straightforward. We can thus improve upon the baseline SSA form, where def/use chains over heap accesses are ignored. Indeed, to ensure feasibility, we must account for such accesses. Hence, we replace accesses to field {\tt f} of local object $o$ with a fresh local variable {\tt l$_{o{\tt .f}}$}, where $o{\tt .f}$ denotes the memory location itself and not its value. This is done prior to the standard SSA transformation, and under the assumption that in the predicted run, the base object $o$ remains the same.

\paragraph{Full Encoding: Method Calls and Shared Accesses}

@@

To account for method invocations, we induce additional context on variable accesses. Instead of recording only the version of a variable, we also represent as part of the variable's identifier its enclosing method and its invocation counter. Thus, symbol ${\tt m}^2 \colon {\tt v}^7$ is interpreted as the definition of local variable ${\tt v}$ within the second invocation of method ${\tt m}$ that occurs at trace index 7.

@@

The final aspect of our encoding process is representation of shared heap accesses. As illustrated in Section \ref{Se:techoverview}, this is done via designated symbols:
\begin{itemize}
	\item $W^{k}_{\tt v}$ denotes write access to variable ${\tt v}$ at trace index $k$ (where ${\tt v}$ is assumed to be a qualified identifier to account for the interprocedural setting).
	\item Analogously, $R^{k}_{\tt v}$ denotes that ${\tt v}$ is read at index $k$.
\end{itemize}

@@

\subsection{Constraint System}

Having explained how the trace is encoded, we now describe in detail how constraints are derived from the trace, such that any permutation considered by the analysis is guaranteed to represent a feasible execution schedule.

\paragraph{Intra-thread Order}

The first set of constraints reflects control flow within the individual threads, which must remain unchanged under the reordering transformation. Given input trace $t$, this is expressed as the following formula:
$$
\begin{array}{rl}
\forall \tau,\tau' \in t. & {\sf thread}\ \tau \equiv {\sf thread}\ \tau'. \\
& {\sf index}\ \tau < {\sf index}\ \tau' \Rightarrow O_{\tau} < O_{\tau'}
\end{array}
$$ 
The logical variables $O_x$ express ordering constraints. The requirement, as stated above, is that these variables
reflect the same order as the projection of ${\sf index}$ onto individual threads.

\paragraph{Race Condition}

Given pair $\tau$ and $\tau'$ of events that both access a common memory location $\ell$, we demand that
$$
\begin{array}{rl}
& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
\bigwedge 	& (\ell \in {\sf writeset}\ \tau \vee \ell \in {\sf writeset}\ \tau') \\
\bigwedge   & O_{\tau} = O_{\tau'}
\end{array}
$$
That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$ (as judged by the ${\sf writeset}$ membership check), and (iii) the events occur simultanesouly. This is a direct logical encoding of the definition of a race condition.

\paragraph{Path Constraints}

The requirement with respect to branching is that the prefix of the original trace $t$ up to the pair $\tau$ and $\tau'$ of candidate racing events remains identical in the predicted trace $t'$. More accurately, under the assumption that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
we require that 
$$
\bigwedge_{\tau'' \in t \cap {\bf bexp}.\
	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
$$ 
where ${\sf stmt}$ is a helper function that obtains the code statement incident in a given transition. That is, all branching transitions up to $\tau'$ (which occurs after $\tau$) preserve their boolean interpretation under $t'$. This ensures that there are no divergences from the path containing the racing events, though beyond that path any feasible continuation is permitted. Importantly, contrary to \cite{JEFF-PLDI14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions is invariant under the input and predicted traces.

\paragraph{Variable Definitions}

A final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before. That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$. This is enforced as the formula
$$
\bigwedge_{\tau'' \in t \cap {\bf asgn}.\
	\tau'' \in {\sf pre}\ \tau'} {\sf stmt}\ \tau'' \in t' 
$$
where we again assume that ${\sf index}\ \tau < {\sf index}\ \tau'$. That is, the same statement occurring in $t$ is also present in $t'$ (though the transitions may differ). Since the statements of $t'$ are a permutation of the statements of $t$, we are assured that use/def flow is constrained appropriately.

\paragraph{Relaxation of Flow Dependencies}

We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the data flow exhibited in the original trace. More precisely, \tool\ is able to relax flow dependencies in the original trace, whereby a thread reads a shared memory location written by another thread, while enforcing feasibility. This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same write access as in the original trace.

To ensure feasibility under relaxation of flow dependencies, we need to secure the link between the execution schedule and the write/read flow. As an illustration from the example in Figure \ref{fig:running}, $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule where thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}, the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}: $R_{\tt y}^4={\tt y}^1$.

The full and general constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
$$
\begin{array}{rll}
\bigvee_{e_w \in {\cal W}} &  & (R_{\ell} = {\ell}^{{\sf index}\ e_w}) \\
&		\bigwedge 	&  O(e_w) < O(e) \\
&		\bigwedge_{e' \in {\cal W} \setminus \{ e_w \}} & (O(e') < O(e_w) \vee O(e) < O(e'))
\end{array}
$$
This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O(e_w) < O(e)$) and (ii) all other write events either occur before ($O(e') < O(e_w)$) it or after the read event
($O(e) < O(e')$).

An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning. As an illustration, we refer to Figure \ref{fig:heapAccess}. While the read at line {\tt 7} appears to match the write at line {\tt 5}, this is conditioned on the read at line {\tt 6} being linked to the assignment at line {\tt 4}. If the predicted run violates this link, then feasibility is no longer guaranteed. In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the write at line {\tt 5} no longer matches the read at line {\tt 7}.

\begin{figure}
	\centering
	\begin{tabular}{ll}
		\hline
		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
		\hline
		{\tt 1: x1 = new();} & \\
		{\tt 2: x2 = new();} & \\
		{\tt 3: y = x1;} & \\
		{\tt 4: y = x2;} & \\
		{\tt 5: x2.f = 5;} & \\	
		& {\tt 6: z = y;} \\
		& {\tt 7: w = z.f;} \\
	\end{tabular}
	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation}
\end{figure}

To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced before the candidate racing events retain their original address in the predicted trace.
This achieves three guarantees: First, matching events in the original trace are guaranteed to also match in the predicted trace. Second, candidate races in the original trace remain viable in the predicted trace. Finally, in a practical setting involving virtual method calls (which goes beyond our core grammar in Table \ref{Ta:syntax}), call-site resolutions are the same across the original and predicted traces. Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.

Formally, given pair $\tau$ and $\tau'$ of candidate racing events such that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
we require that
$$
\bigwedge_{\tau'' \equiv {\tt y=x.f} \in t \cap {\bf heapr}.\
	\tau'' \in {\sf pre}\ \tau'} {\sf env}\ \sigma(t,\tau'')\ {\tt x} = {\sf env}\ \sigma(t',\tau'')\ {\tt x}
$$
where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
This constraint fixes that all heap dereferences up to the later of the candidate racing events retain their original base object as in the predicted trace $t'$.

\subsection{Correctness Proof}

{\bf MISSING!}
